Dark energy is based on a math error from 1930

### 2024-12-26 paper/paper.tex
Add a code example

\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amsmath}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{csquotes}

\title{Dark energy is due to a math error from 1930}

@@

\hypersetup{
pdftitle={Dark energy is due to a math error from 1930},
pdfsubject={astro-ph.CO},
pdfauthor={Logan P.~Evans},
pdfkeywords={cosmological parameters, dark energy},
}

@@

\begin{abstract}
  Light from distant objects, such as supernovae and galaxies, is dimmed
  due to three phenomena associated with redshift: energy levels per photon,
  spectral bandwidth, and time dilation. Observed magnitudes are reported with
  k-corrections that should correct for these dimming effects, but since 1930,
  the equations used to perform these k-corrections have been missing the
  correction factor for time dilation. When observations of Type Ia supernova
  are corrected for this missing factor, the relationship between luminosity
  distance and redshift for Type Ia supernova becomes linear. This indicates
  that the expansion rate of the universe is not accelerating and there is no
  need for dark energy to explain observational data.
\end{abstract}

% keywords can be removed
\keywords{Cosmological Parameters \and Dark Energy \and Luminosity Distance}

@@

Observations of Type Ia supernova are fundamental to the study of cosmology.
These measurements are necessary to calibrate models that measure redshift and
estimate distance. This relationship, often called the Hubble-Lema\^{i}tre law,
[TODO: find citation] describes how quickly the universe is expanding. However,
\citet{riess1998} and \citet{perlmutter1999} presented evidence that there
isn't a linear relationship between redshift and distance, but instead, distant
objects are farther away than their redshift would predict (see Figure
\ref{fig:mu_distance_vs_redshift}). This phenomenon is referred to as dark
energy.

@@

It turns out that measuring the apperant brightness of a Type Ia supernova is
non-trivial. The aspect of the process that we will explore here concerns how
redshift affects the light we observe. This problem is often referred to as
k-corrections, and one of the first mathematical treatments of the problem was
performed by \citet{tolman1930}. However, when Tolman made his derivation, he
did not consider the effects of a spectra that is stretched out due to
redshift.

@@

By \citet{oke1968}, the two factors of $(1 + z)$ were attributed to the change
in energy and to the spectral bandwidth elongation, which leaves time dilation
as the factor that was omitted.

In 1935, the correction of $(1 + z)^2$ empirically lined up the data so that
there seemed to be a linear relationship between redshift and distance. It
wasn't until \citet{riess1998}, sixty-eight years later, that the omission of a
third $(1 + z)$ correction factor resulted in faulty conclusions.

Those faulty conclusions, however, have been the basis for a substantial amount
of scientific interest. In 2011, three physicists received the Nobel Prize in
Physics for discovering that the expansion rate of the universe is accelerating
\citet{straumann2012}. The papers \citet{riess1998} and \citet{perlmutter1999}
have each received over 20000 citations. It's not clear how much cosmology work
will need to be revisited with corrected distance measurements, but it is a
non-trivial amount.

\begin{displayquote}
\end{displayquote}

\section{The dimming effects of redshift}

@@

\section{Correcting magnitude for time dilation}

@@

\section{Linear distance vs redshift relationship}

@@

\begin{figure}[ht]
  \includegraphics[width=\linewidth]{mu_distance_vs_redshift.png}
  \caption{The relationship between distance and redshift for two treatments of
  magnitude data. The displayed points are roughly a third of the values in the
  full DES dataset, selected evenly to aid visibility. The $k(z) = 1$ treatment
  is clearly non-linear while the $k(z) = 1 + z$ treatment appears to be
  linear.}
  \label{fig:mu_distance_vs_redshift}
\end{figure}

\begin{figure}[ht]
  \includegraphics[width=\linewidth]{k-corrections_for_photon_counts.png}
  \caption{An example of how k-corrections take an observed spectrum and
  produce a rest-frame spectrum. [TODO]
  }
\end{figure}

@@

@@

\section{Disagreement with existing research}

>>>>

\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amsmath}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{csquotes}
\usepackage{float}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{fancyvrb}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=left,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2
}

\lstset{style=mystyle}

\title{Dark energy is based on a math error from 1930}

@@

\hypersetup{
pdftitle={Dark energy is based on a math error from 1930},
pdfsubject={astro-ph.CO},
pdfauthor={Logan P.~Evans},
pdfkeywords={cosmological parameters, dark energy},
}

@@

\begin{abstract}
  We show that the magnitude of distant objects has been calculated incorrectly
  since 1930. We describe how this math error has propagated for over nine
  decades. Finally, we explore the consequences of correcting this flaw and
  resolve two of the preeminent mysteries in cosmology: dark energy is not
  supported by observations of Type Ia supernovae, and the Hubble tension is
  due to a calculation error.
\end{abstract}

% keywords can be removed
\keywords{Cosmological Parameters \and Dark Energy \and Luminosity Distance \and Hubble tension}

@@

The measurement of Type Ia supernovae is one of the primary sources of data for
cosmological models. The measurements involve estimating the peak magnitude in
the B filter \citet{riess1998} for the rest frame, meaning we want to know how
bright the SnIa would be if its light didn't undergo redshift. Redshift causes
the light that would be mesaured by the B filter in a rest frame to be shifted
to another filter. In order to observe a SnIa in an arbitrary filter and report
the magnitude in filter B, we use a technique called k-corrections.

The first formal derivation of k-corrections was performed by
\citet{tollman1930}, and it failed to account for bandwidth stretching, one of
the dimming effects of redshift. This error was noted by \citet{desitter1934},
but in \citet{hubble1935} the descrepancy was noted and ignored.
\citet{oke1968} rederived k-corrections, but they failed to account time
dilation, one of the other dimming effects of redshift. In \citet{kim1996},
k-corrections were extended to handle additional cross-filter comparisons and
address zero-point corrections, but they started their derivation by
referencing the incorrect equation from \citet{oke1968}. \citet{hsiao2007}
expanded on the work done by \citet{kim1996} by providing expected SnIa
template spectra.  Finally, \citet{burns2010} implemented the popular software
package Sn(OO)py which implemented the SnIa template spectra into
k-corrections. This software package computes k-corrections differently from
prior work, and in order to match prior values, the code adds an extra term.
This history is explored in more depth in Section \ref{sec:history}.

The effect of using an incorrect magnitude for SnIa is that we think objects
are farther away than they really are, and the effect componds for greater
distances. Up until \citet{riess1998}, we didn't have distant enough
observations for this error to matter much for cosmology models, but in the
late 1990s, it was clear that our measurements for distance and redshift were
not linear. This observation led to a model that utilized a cosmological
constant and dark energy in order to explain the non-linear distance-redshift
graph. We will show in Section \ref{sec:correction} that correcting the flaw
with k-corrections leads to a linear graph that does not need to rely on a
cosmological constant.

Since \citet{planck2020}, it has been evident that something was missing with
our understanding of cosmology. This research used an alternative technique
based on measurements of the cosmic microwave background to measure the Hubble
constant, but as summarized in \citet{perivolaropoulos2022}, this measurement
was incompatable with the Hubble constant measured from SnIa. As we will
discuss in Section \ref{sec:tension}, fixing the error with k-corrections
produces a measurement of the Hubble constant that is compatible with models
based on the cosmic microwave background.

\section{A brief history of k-corrections}
\label{sec:history}

Observations of Type Ia supernova are essential to calibrate models that
measure redshift and estimate distance. This relationship, often called the
Hubble-Lema\^{i}tre law, describes how quickly the universe is expanding.
However, \citet{riess1998} and \citet{perlmutter1999} presented evidence that
there isn't a linear relationship between redshift and distance, but instead,
distant objects are farther away than their redshift would predict (see Figure
\ref{fig:mu_distance_vs_redshift}). This phenomenon implies that the
acceleration of the universe is faster today than it was for old observations.
Previously the cause of this phenomenon was unknown and was referred to as
dark energy.

\begin{figure}[h]
  \includegraphics[width=\linewidth]{mu_distance_vs_redshift.png}
  \caption{The relationship between distance and redshift for two treatments of
  magnitude data. The displayed points are roughly a third of the values in the
  full DES dataset, selected evenly to aid visibility. The $k(z) = 1$ treatment
  is clearly non-linear while the $k(z) = 1 + z$ treatment appears to be
  linear.}
  \label{fig:mu_distance_vs_redshift}
\end{figure}

@@

Measuring the apperant brightness of a Type Ia supernova is non-trivial. The
aspect of the process that we will explore here concerns how redshift affects
the light we observe. This problem is often referred to as k-corrections, and
one of the first mathematical treatments of the problem was performed by
\citet{tolman1930}. However, when Tolman made his derivation, he did not
consider the effects of a spectra that is stretched out due to redshift.

@@

By \citet{oke1968}, the two factors of $(1 + z)$ were attributed to the change
in energy and to the spectral bandwidth elongation, which leaves time dilation
as the factor that was omitted. A graph that demonstrates why it's essential to
correct for both the spectra bandwidth warping and time dilation is presented
in Figure \ref{fit:k-example}.

\begin{figure}[h]
  \includegraphics[width=\linewidth]{k-corrections_for_photon_counts.png}
  \caption{An example of how k-corrections take an observed spectrum and
  produce a rest-frame spectrum. [TODO]
  }
  \label{fig:k-example}
\end{figure}

The modern treatment of k-corrections is based on the work of \citet{kim1996}.

\lstinputlisting[language=Python, caption=A demonstration that k-corrections
calculated by snpy are off by \texttt{2.5*log10(1+z)}. The roundoff error is
roughly \texttt{2e-16}.]{code_listing.py}

\section{The dimming effects of redshift}
\label{sec:redshift}

@@

\section{Correcting magnitude for time dilation}
\label{sec:correction}

@@

\section{Linear distance vs redshift relationship}
\label{hubblelaw}

@@

@@

\section{Hubble tension}
\label{sec:tension}

@@
